Reasoning Graphically about Quantum Computation

### 2008-03-07 papers/reasoning-graphically/paper.tex
bibtex cleaning and squeesing to 15 pages.
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1409 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

\textbf{Matching.} To formalise the intuitive notion that a !-box
represents arbitrary number of copies of the subgraph made from it's
nodes, we introduce \emph{!-box matching}. This is a binary relation
on graphs, written infix as $G \leq_! H$ when $G$ matches $H$. The set
matching !-box graphs is defined as the set closed by the following
operations on the !-boxes in a graph:

@@

\textbf{Semantics.} We give a formal semantics to a !-box graphs by
defining the !-box graph in terms of a set of graphs in the underlying
representation. We denote the interpretation of a !-box graph $G$ by
$\binterp{G}$ and say that members of $\binterp{G}$ are instances of
$G$.

@@

\textbf{Semantics as Concrete Matches.} The semantics for a graph
pattern $G$ is a set of concrete graphs denoted by $\minterp{G}$ and
define it as:

@@

\textbf{Graph Pattern Matches.} The specification for one pattern,
$G$, to match another one, $H$ is that it is more general with respect
to the interpretation: $\minterp{G} \supseteq \minterp{H}$. However,
graph patterns can correspond to a countably infinite number of
concrete graphs. Thus matching between graph pattern cannot be
implemented by simply unfolding all interpretions as concrete graphs
and checking the membership relation.

@@

\textbf{Computing the Quantum Fourier Transform.} The quantum Fourier
transform is among the most important quantum algorithms, forming an
essential part of Shor's algorithm~\cite{Shor:PolyTimeFact:1997},
famous for providing a polynomial time algorithm for factoring. In our
graph patten calculus this circuit becomes the top-left graph in
Figure~\ref{fig:quantum-transform}. This figure shows how computation
can be symbolically performed by rewriting with the lifted equations
from Figure~\ref{fig:graph-quant-eqns} and the graph pattern version
of the Spider Theorem.

>>>>

%\textbf{Matching.} 
To formalise the intuitive notion that a !-box
represents arbitrary number of copies of the subgraph made from it's
nodes, we introduce \emph{!-box matching}. This is a binary relation
on graphs, written infix as $G \leq_! H$ when $G$ matches $H$. The set
matching !-box graphs is defined as the set closed by the following
operations on the !-boxes in a graph:

@@

%\textbf{Semantics.} 
We give a formal semantics to a !-box graphs by
defining the !-box graph in terms of a set of graphs in the underlying
representation. We denote the interpretation of a !-box graph $G$ by
$\binterp{G}$ and say that members of $\binterp{G}$ are instances of
$G$.

@@

%\textbf{Semantics as Concrete Matches.} 
The semantics for a graph
pattern $G$ is a set of concrete graphs denoted by $\minterp{G}$ and
define it as:

@@

%\textbf{Graph Pattern Matches.} 
The specification for one pattern,
$G$, to match another one, $H$ is that it is more general with respect
to the interpretation: $\minterp{G} \supseteq \minterp{H}$. However,
graph patterns can correspond to a countably infinite number of
concrete graphs. Thus matching between graph pattern cannot be
implemented by simply unfolding all interpretions as concrete graphs
and checking the membership relation.

@@

The quantum Fourier transform is among the most important quantum
algorithms, forming an essential part of Shor's
algorithm~\cite{Shor:PolyTimeFact:1997}, famous for providing a
polynomial time algorithm for factoring. In our graph patten calculus
this circuit becomes the top-left graph in
Figure~\ref{fig:quantum-transform}. This figure shows how computation
can be symbolically performed by rewriting with the lifted equations
from Figure~\ref{fig:graph-quant-eqns} and the graph pattern version
of the Spider Theorem.
